### 2025-04-27 thesis/thesis.tex
Write a chapter about limitations

\citeauthor*{SQLizer} show that alternating between quantitative repair and synthesis yields results that beat
state-of-the-art approaches like NaLIR. SQLizer is fully automated and database-agnostic, requiring no knowledge
of the underlying schema. The authors evaluated SQLizer on 455 queries across three databases, where it ranked the
correct query in the top 5 results for about 90\% of cases, significantly outperforming NaLIR \citep{NALIR},
the previous state-of-the-art system \citep{SQLizer}.

@@

>>>>

\citeauthor*{SQLizer} show that alternating between repairing and synthesis yields results that beat state-of-the-art
NL2SQL approaches like NaLIR. SQLizer is fully automated and database-agnostic, requiring no knowledge of the underlying
schema. The authors evaluated SQLizer on 455 queries across three databases, where it ranked the correct query in the
top 5 results for roughly 90\% of the queries. This represents a significant improvement over NaLIR \citep{NALIR},
the previous state-of-the-art system \citep{SQLizer}.

@@

\subsubsection{Limitations of Traditional Approaches to NL2SQL}

Despite being innovative and achieving state-of-the-art results, many of the above outlined approaches face severe 
challenges when moving outside of an research environment. Many of these systems performed comparatively good on research
benchmarks that were often composed of controlled question types and limited data variety. Ultimatively no standard
benchmark existed for NL2SQL in this era, hence comparing different NL2SQL systems against each other is a problem on 
its own. Despite not having a standard benchmark that all approaches could be unifiably evaluated against, several 
fundamental challenges emerged / remained with these approaches:

\begin{enumerate}
    \item \textbf{Limited linguistic coverage} — Prevalent rule-based and semantic-parsing based systems were only able to 
          process the a small subset of the natural language they were programmed for. This severely limited their 
          ability to handle different phrasings of the same end-user goal \citep{StringKernels, UnnaturalQueryLanguage, 
          Lunar, Ladder}.
    \item \textbf{Transferability} — Traditional approaches typically required extensive manual configuration or at least
          a training phase / adaption for each database they were deployed for, hindering cross domain usage through being 
          expensive and time-consuming to adapt \citep{NLIDBs, Lunar}.
    \item \textbf{Brittleness} — Many of the systems introduced in this subchapter did not handle synonyms, paraphrasing,   
          or spelling errors well. Manual adaption / handling was needed in order to becomes resilient against each class
          of problems \citep{StringKernels, SQLizer}.
    \item \textbf{Poor scalability} — With potentially more complex underlying databases, traditional 
          solutions often showed to perform worse. \citeauthor*{GraphMatching} found, that with increasing schema
          complexity more compute was required to resolve the natural language query to a suitable query candidate
          making them less transferable and scalable than initially anticipated \citep{GraphMatching} — ``Evaluating on
          all domains in Freebase would generate a very large number of queries for which denotations would have to be 
          computed ... Our system loads Freebase using Virtuoso and queries it with SPARQL. Virtuoso is slow in dealing with
          millions of queries indexed on the entire Freebase, and is the only reason we did not work with the complete
          Freebase.'' which indicates underlying system design issues with runtime complexity.
\end{enumerate}

These flaws of traditional NL2SQL approaches made it apparent, that a different class of approaches is needed, which 
increase transferability and reduce the brittleness since users are ``unwilling to trade reliable and predictable user 
interfaces for intelligent but unreliable ones'' according to \cite{NLIDBTheory}. Whilst many approaches outlined tractable 
ways to increase user satisfaction and accuracy (like \citeauthor*{Rendezvous} did in \citeyear{Rendezvous} with a 
conversational approach), NLIDBs were and are not considered to be a solved problem.
